---
title: "data visualization with ggplot2"
author: "Julia Silge"
email: "@juliasilge"
ratio: 16x10
maketitle: true
output:
  rmdshower::shower_presentation:
    self_contained: false
    katex: true
    theme: stack
---

## Understanding the grammar of graphics

- ggplot2 is a data visualization library for R
- Developed by Hadley Wickham
- ggplot2 is based on a **grammar**
- It's powerful, flexible, beautiful, based on the features in your data, designed for iterative workflows

# Do not draw with your data

## Elements of the grammar

> - Data + mapping (aesthetic)
> - Layers
> - Scales
> - Coordinate systems
> - Faceting
> - Themes


## Gapminder data

```{r}
library(gapminder)
gapminder
```

## Data + aesthetic mapping

```{r, fig.width=4.5, fig.height=4.5}
library(ggplot2)
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp))
```

## Now add a layer

```{r, fig.width=4.5, fig.height=4.5}
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp)) +
    geom_point()
```

## Now add a scale

```{r, fig.width=4.5, fig.height=4.5}
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp)) +
    geom_point() +
    scale_x_log10()
```


## Another aesthetic mapping?

```{r, fig.width=6, fig.height=4.5}
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp, color = continent)) +
    geom_point() +
    scale_x_log10()
```

## Too many points `r emo::ji("thinking")`

```{r, fig.width=6, fig.height=4.5}
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp, color = continent)) +
    geom_point(alpha = 0.4, size = 2) +
    scale_x_log10()
```

## Add a best fit line `r emo::ji("smiley")`

```{r, fig.width=6, fig.height=4.5}
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp, color = continent)) +
    geom_point(alpha = 0.4, size = 2) +
    scale_x_log10() + geom_smooth(method = "lm")
```


## Speaking of lines...

```{r, fig.width=6, fig.height=4.5}
ggplot(gapminder, aes(x = year, y = lifeExp, color = country)) +
    geom_line(size = 1.5, alpha = 0.8, show.legend = FALSE) +
    scale_color_manual(values = country_colors)
```

## Facetting

```{r, fig.width=9, fig.height=3}
ggplot(gapminder, aes(x = year, y = lifeExp, color = country)) +
    geom_line(size = 1.5, alpha = 0.8, show.legend = FALSE) +
    scale_color_manual(values = country_colors) +
    facet_wrap(~continent, nrow = 1)
```

## Themes

```{r, fig.width=9, fig.height=3}
ggplot(gapminder, aes(x = year, y = lifeExp, color = country)) +
    geom_line(size = 1.5, alpha = 0.8, show.legend = FALSE) +
    scale_color_manual(values = country_colors) +
    facet_wrap(~continent, nrow = 1) +
    theme_minimal()
```

## Themes

```{r, fig.width=9, fig.height=3}
library(silgelib)
ggplot(gapminder, aes(x = year, y = lifeExp, color = country)) +
    geom_line(size = 1.5, alpha = 0.8, show.legend = FALSE) +
    scale_color_manual(values = country_colors) +
    facet_wrap(~continent, nrow = 1) +
    theme_roboto()
```

# Tidy data principles = `r emo::ji("tada")`

## More resources

- About ggplot2: http://ggplot2.tidyverse.org/
- Extensions to ggplot2: http://www.ggplot2-exts.org/
- [Data visualization chapter](http://r4ds.had.co.nz/data-visualisation.html) in *R for Data Science*
- [*ggplot2: Elegant Graphics for Data Analysis*](http://amzn.to/2fncG50)

